---
title: "Homework 3"
author: "Kylie Wheelock Riley"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 
```{r importing problem 1 data, include = FALSE}
insta_data = readr::read_csv(file = "./data/instacart.csv.zip") 

##The dataset has 1,384,617 rows of data for 131,206 distinct orders
nrow(insta_data)
n_distinct(pull(insta_data, order_id))


## Looking at individual variables 
#order_id - order ID has a mean of 1706298 and max value of 3421070
mean(pull(insta_data, order_id))
max(pull(insta_data, order_id))

insta_data %>% 
  group_by(order_id) %>% 
  summarize(
    n_obs = n(),
    n_distinct(product_id)
  ) 
 

#product_id - there are 39,123 unique products ordered in this dataset, and product 24852 is the most popular with 18,726 orders. 
n_distinct(pull(insta_data, product_id))

insta_data %>% 
  group_by(product_id) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) 
  

#add_to_cart_order - the maximum add to cart order was 80
max(pull(insta_data, add_to_cart_order))
 


#reordered - 60% of items were reorders and 40% were new. Plot showing this. ## Want to figure out how to plot a percent. 
insta_data %>% 
  group_by(reordered) %>%
  summarize(count = n()) %>%
  mutate(
    percent = round((count/sum(count)*100))
    )


ggplot(insta_data, aes(x = reordered)) + 
  geom_histogram(binwidth = 0.45) +
  labs(
    title = "frequency of items that are reordered",
    x = "Reorder",
    y = "Frequency",
    caption = "Instacart grocery order data."
  ) +
  scale_x_continuous(
    breaks = c(0, 1), 
    labels = c("no", "yes")
    )


#order_hour_of_day - 2 pm is the most pupular time of day to place food orders. 
insta_data %>% 
  group_by(order_hour_of_day) %>%
  summarize(count = n()) %>%
  arrange(desc(count))


#days_since_prior_order - on average people wait 17 days to place their next food order. 
mean(pull(insta_data, days_since_prior_order))


#product_name - produces a table with the top 10 items ordered.  The number 1 ordered item was bananas. 
insta_data %>% 
  group_by(product_name) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>% 
  slice(1:10) %>% 
  knitr::kable()

```

**1A) How many aisles are there, and which aisles are the most items ordered from?**
```{r problem 1 aisle count, include = FALSE}
max(pull(insta_data, aisle_id))
n_distinct(pull(insta_data, aisle_id))
```
There are `r n_distinct(pull(insta_data, aisle_id))` aisles at the Instacart warehouse.  The most popular aisle is Fresh Vegetables, and the top 5 aisles are listed below. 
```{r most popular aisle table, echo = FALSE}
insta_data %>% 
  group_by(aisle) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>% 
  slice(1:5) %>%
  knitr::kable() 
```


**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}

```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}

```

